### 2010-04-30 trunk/htdocs/sections.tex
Added notes on polar active gel from Davide
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@993 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  sections.tex
%
%  Content for Ludwig documentation.
%
%  $Id: sections.tex,v 1.12 2010-01-21 18:20:51 kevin Exp $
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%  (c) 2008 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  sections.tex
%
%  Content for Ludwig documentation.
%
%  $Id: sections.tex,v 1.14 2010-04-30 11:00:50 kevin Exp $
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%  (c) 2008 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{Polar Active Gel}

Here we consider a polar vector field model to study active gels. 
The order parameter is this time a vector $P_\alpha$
(with variable magnitude) as there is no longer head-tail symmetry
in the system as was the case for the ${\bf Q}$ tensor formulation. 
The equation of motion as:
\begin{equation}
\partial_t P_\alpha  \partial_\beta (u_{\beta} + wP_{\beta}) P_{\alpha}
= \lambda D_{\alpha\beta} P_{\beta} - \omega_{\alpha\beta} P_{\beta}
+\Gamma' h_{\alpha}.
\end{equation}

In this equation, $w$ is an active term, due to swimming, which causes
self-advection of the order parameter, while $\lambda$ is a material
dependent constant -- positive for rod-like molecules. If $|\lambda|>1$
the liquid crystalline passive phase is flow-aligning, otherwise it is
flow-tumbling.

Note that the convection for the Leslie-Ericksen is that the
velocity gradient tensor should be writen $W_{\alpha\beta}
= \partial_\alpha u_\beta$ (and not the other way round, which
is usual).

The molecular field is given by
$h_{\alpha}=-\delta {\cal F}_{\rm pol}/\delta p_{\alpha}$
where ${\cal F}_{\rm pol}$ is the free energy for a polar active nematic,
whose density we can take to be:
\begin{equation}
f =  \frac{\alpha_2}{2} |{\mathbf P}|^2 + \frac{\alpha_4}{4}|{\mathbf P}|^4
+ \frac{K}{2}\left(\partial_\alpha P_{\beta}\right)^2 + 
\frac{K_{LC}}{2}\left[\partial_\alpha (P_{\beta}P_{\gamma})\right]^2 
\end{equation}
where $K$ and $K_{LC}$ are elastic constants, $\alpha_4>0$, and $\alpha_2$ 
can be positive (isotropic phase) or negative (polar nematic phase).

The molecular field is explicitly given by:
\begin{equation}
h_{\alpha}=-\alpha_2 P_{\alpha} -\alpha_4 P_{\alpha} P_{\beta}P_{\beta}+
K \partial_{\beta}\partial_{\beta} P_{\alpha}+2 K_{LC}
P_{\gamma}\partial_{\beta}\partial_{\beta} (P_{\gamma}P_{\alpha}).
\end{equation}

The Navier-Stokes equation is:
\begin{equation}
\rho \left(\partial_t+v_{\beta}\partial_{\beta}\right) v_{\alpha} =
\partial_{\beta} \Pi_{\alpha\beta} 
+\eta \partial_{\beta}\partial_{\beta} v_{\alpha}.
\end{equation}

The stress tensor this time is 
\begin{eqnarray}
\Pi_{\alpha\beta} =
\frac{1}{2}\left(P_\alpha h_\beta - P_\beta h_\alpha \right)
- \lambda\left( (P_\alpha h_\beta + P_\beta h_\alpha)/2 
- (1/3)P_\gamma P_\gamma \delta_{\alpha\beta} \right)
\\
-  \zeta ( P_\alpha P_\beta - (1/3)P_\gamma P_\gamma \delta_{\alpha\beta})
- K \partial_\alpha P_\gamma \partial_\beta P_\gamma.
\end{eqnarray}
The active term in the stress tensor, proportional
to $\zeta$, has the same form for apolar and polar active gels.

@@

\subsection{D2Q9}

These are the eigenvalues and eigenvectors of the collision
matrix used for D2Q9

\begin{table}
\begin{tabular}{|l|r|rrrrrrrrr|r|l|}
\hline\hline
$M^a$ & & \multicolumn{9}{c|}{$m_i^a$} & $N^a$  &\\
\hline
$\rho$ & - & 1 &  1 &  1 &  1 &  1 &  1 &  1 &   1 &  1 & 1 &$\mathbf{1}$ \\
\hline
$\rho c_{ix}$ & - & 0 &  1 &  1 & 1 & 0 &  0 & -1 &  1 & -1 & 3 & $c_{ix}$ \\
\hline
$\rho c_{iy}$ & - & 0 & 1 &  0 &  -1 &  1 &  -1 & 1 & 0 & -1 & 3  &$c_{iy}$ \\
\hline
$Q_{xx}$ & 1/3 & -1 &  2 &  2 & 2 & -1 & -1 & 2 & 2 & 2 & 9/2 
& $c_{ix} c_{ix} - c_s^2$ \\
\hline
$Q_{xy}$ & - & 0 &  1 & 0 & -1 & 0 & 0 & -1 & 0 & 1 & 9 & $c_{ix} c_{iy}$ \\
\hline
$Q_{yy}$ & 1/3 & -1 &  2 & -1 & 2 & 2 & 2 & 2 & -1 & 2 & 9/2
& $c_{iy} c_{iy} - c_s^2$ \\
\hline\hline
$\chi^1$ & - &  1 & 4 & -2 & 4 & -2 & -2 & 4 & -2 & 4 & 1/4 & $\chi^1$ \\
\hline
$J_{ix}$ & - & 0 &  4 & 0 & -4 & -2 & 2 & 4 & 0 & -4 & 3/8
& $\chi^1 \rho c_{ix}$\\
\hline
$J_{iy}$ & - & 0 & 4 & 0 & -4 & -2 & 2 & 4 & 0 & -4 & 3/8
& $\chi^1 \rho c_{iy}$\\
\hline\hline
$w_i$ & 1/36 & 16 & 1 & 4 & 1 & 4 & 4 & 1 & 4 & 1 & & $w_i$\\
\hline\hline
\end{tabular}
\end{table}
